---
title: "Hong_1992"
author: "Hong Guo"
date: "8/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Hi Hong!
Greetings from my laptop!!
Bob  

Are you ready for a break?
```{r}
suppressPackageStartupMessages({
  library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
})
```

# Load salmon data from KNB

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92020.1"
# data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"
esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)
# First try to get local copy or find it in the website
head(esc)
```
# Prepare the dataset for calculating medions

```{r}
median_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  group_by(Species, SASAP.Region, Year, Location) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  group_by(Species) %>% 
  summarize(median_escapement = median(escapement))

head(median_esc)
```

# Plot the escapement as a bar chart

```{r}
ggplot(median_esc, aes(Species, median_escapement)) +
  geom_col() +
  coord_flip() 
# geom_col is a bar chart and coord_flip is a convert with xy coordiaate
```
# Now work on building a map

We're starting from daily escapement data again, and finding the unique locations at which escapement was measured.

## Find unique location point 

Make sure to remember to deal with missing values in the location field.

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
  # distinct is a useful function to extract the unique values
datatable(locations)
```

## Draw the map

Use leaflet to make an interactive map.

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
# You could generate your own tile or marker
```

```{r}
locs <- locations %>% mutate(Longitude = abs(Longitude) * -1)

leaflet(locs) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
```

```{r}
# Use a custom marker so Leaflet doesn't try to grab the marker images from 
# its CDN (this was brought up in 
# https://github.com/NCEAS/sasap-training/issues/22)
markerIcon <- makeIcon(
  iconUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/1.3.1/images/marker-icon.png",
  iconWidth = 25, iconHeight = 41,
  iconAnchorX = 12, iconAnchorY = 41,
  shadowUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/1.3.1/images/marker-shadow.png",
  shadowWidth = 41, shadowHeight = 41,
  shadowAnchorX = 13, shadowAnchorY = 41
)
```

```{r}
leaflet(locs) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location, icon = markerIcon)
```

