---
title: "Tidying Data in R"
author: "Hong Guo"
date: "8/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up 

```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(tidyr)
```

Calling a masked function

packageName::functionName
example:

stats::filter()

# Data input

Read in Data from [Mike Byerly. 2016. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1)

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```


the pipe (%>%) operate
command +shift+M for pipe operator
command +shift+N for new R

select filter

```{r}
catch_df <- catch_df %>% 
  select(-All,-notesRegCode) 
# select positively by name
# select(Region,Year,Chinook,Sockeye,Coho,Pink,Chum) 
head(catch_df)
```

Move from wide to long format (using gather)
Reverse of gather is spread
Rename a column


```{r}
catch_df <- catch_df %>% 
  gather(key = "species",value = "catch",-Region,-Year) %>% 
  rename(catch_thousands = catch)
unique(catch_df$species)
```

Create a new column

1) turn "I" into "1" in the catch_thousands column
2) coerce catch_thousands into integer type
3) convert catch_thousands to catch

```{r}
catch_df <- catch_df %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I",1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands *1000)
head(catch_df)
```




