---
title: "Tidying Data in R"
author: "Hong Guo"
date: "8/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up 

```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(tidyr)
```

Calling a masked function

packageName::functionName
example:

stats::filter()

# Data input

Read in Data from [Mike Byerly. 2016. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1)

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```


the pipe (%>%) operate
command +shift+M for pipe operator
command +shift+N for new R

select filter

```{r}
catch_df <- catch_df %>% 
  select(-All,-notesRegCode) 
# select positively by name
# select(Region,Year,Chinook,Sockeye,Coho,Pink,Chum) 
head(catch_df)
```

Move from wide to long format (using gather)
Reverse of gather is spread
Rename a column


```{r}
catch_df <- catch_df %>% 
  gather(key = "species",value = "catch",-Region,-Year) %>% 
  rename(catch_thousands = catch)
unique(catch_df$species)
```

Create a new column

1) turn "I" into "1" in the catch_thousands column
2) coerce catch_thousands into integer type
3) convert catch_thousands to catch

```{r}
catch_df <- catch_df %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I",1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands *1000)
tail(catch_df)

```

# Split-Apply-Combines

```{r}
catch_summarized <- catch_df %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch))
head(catch_summarized)
```

```{r}
catch_year <- catch_df %>% 
  filter(Year >1990 | Year < 1900) %>% 
  group_by(Year,species) %>% 
  summarise(catch_sum = sum(catch))
head(catch_year)
```


```{r}
catch_range <- catch_df %>% 
  group_by(Region) %>% 
  summarise(catch_low = range(catch)[1],
            catch_high = range(catch)[2])

head(catch_range)
```

```{r}
catch_me_sd <- catch_df %>% 
  group_by(Region, species) %>% 
  # calculate mean, sd, n
  summarise(catch_mean = mean(catch),
            catch_sd = sd(catch),
            num_obs = n()) %>% 
  # filter for Chinook
  filter(species == "Chinook" | species == "Chum" ) 
  # | means or; & means and.

head(catch_me_sd)
```

notice: mutate a column for decade calculation

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)

region_defs <- region_defs %>% 
  select(code,mgmtArea)

catch_joined <- left_join(catch_df,region_defs, by = c("Region" = "code"))
# c( ) means the left table is equal to right table

new_catch_df <- catch_df %>% 
  group_by(Region) %>% 
  summarise(total_catch = sum(catch)) %>% 
  left_join(region_defs, by = c("Region" = "code"))


```

